---
title: "Using GPE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using GPE}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, dpi = 200)
```

## Prerequisites

You'll need tabular data with addresses, or "lon" and "lat" columns containing WG84 (Mercator) location coordinates. Examples are a table with the addresses of schools that program beneficiaries attend, or a street or centroid identifying their neighborhoods, or their home addresses if using personal data is allowed.

## Geocoding

You can geocode your data (obtain latitude and longitude from addresses) using `GPE_geocode()`, a function that queries the Google Maps Platform to translate addresses into latitude/longitude coordinates. 

Using the Google Maps Platform requires a registered API key. To obtain it, follow instructions from  https://developers.google.com/maps/gmp-get-started. Make sure you enable the Geocoding API.

A valid API key is a string of letters and numbers, for example `AIjaSyBR76W62lloYPh_c01LYGhCOZuKU6RVW9` -this is not a real ID by the way, just an example. Once you have your key, you can take a data frame with an address column like this one:

```{r}
library(GPE)

schools
```

...and add latitude and longitude columns using GPE_geocode, specifying the name of the address column and your Google Maps Platform API key:

```{r eval=FALSE}
key <- "AIjaSyBR76W62lloYPh_c01LYGhCOZuKU6RVW9" # This is not a real API key, you must provide yours

GPE_geocode(schools, address, key)
```

```{r echo=FALSE}
key <- ggmap::google_key()

GPE_geocode(schools, address, key)
```

Columns with latitude and longitude coordinates are added to the original data frame.

## Mapping geocoded data

You'll need a data frame containing participant data. 

GPE includes "participants", an example data frame with fictional public program beneficiaries:

```{r}
head(participants)
```


The geographic position of participants and program locations can be plotted, over a basemap, using:

```{r}
GPE_plot_map(participants)
```

In order for `GPE_plot_map()` to work, the input data frame must include columns named "lat" and "lon" representing WGS84 coordinates. As previously shown, latitude and longitude columns can be obtained from addresses using `GPE_geocode()`

In addition, a data frame containing locations -places that participants visit in order to interact with the program, such as training centers, day care providers, etc.- can also be mapped. GPE includes "locations", an example data frame with fictional public program sites:

```{r}
head(locations)
```

```{r}
GPE_plot_map(participants, locations)
```

You can also visualize participant attributes by indicating the name of the column that should be used:

```{r}
GPE_plot_map(participants, locations, participant_attribute = "group")
```

The same can be done for location attributes...

```{r}
GPE_plot_map(participants, locations, location_attribute = "type")
```


... or both participant and location attributes:

```{r}
GPE_plot_map(participants, locations, participant_attribute = "group", location_attribute = "type")
```


## Estimating travel distance and time

Using records of interaction between people and places -representing trips by consumers/beneficiaries to points of sale/access- GPE can estimate time and distance travelled with the Google Maps Platform. As is the case with geocoding, a valid API key is needed to access this service.

GPE provides an example "visits" data frame:

```{r}
visits
```

Trip distance and duration can be obtained using `GPE_travel_time_dist()`. The function takes as input visits data, as well as the locations and participants data frames that provide the origin and destination coordinates. A valid Google Maps Platform API key is also required. Transport mode can be choose from "transit" (default), "driving", "walking", or "bicycling". Keep in mind that transit routing information is not available for all cities; "driving" and "walking" routing is usually available everywhere.   


```{r eval=FALSE}
GPE_travel_time_dist(visits, participants, locations, key)
```

```{r echo=FALSE}
merge(visits, visits_timedist, all = FALSE)
```


## Summaries


GPE includes a simple summary function that takes a "visits" data frame, as described before, and returns basic descriptive statistics for 

 * frecuency of visits, by participant (person) 
 * frecuency of visits, by location (site) 
 
If the input data frame includes `time_minutes` and `distance_km` columns (i.e. as a result of using `GPE_travel_time_dist()`) the summary will also include basic descriptive statistics for 

 * travel time, in minutes 
 * travel distance, in km

For example, given a data frame like

```{r}
head(visits_timedist)
```

the result is:

```{r}
GPE_summary(visits_timedist)
```

## Visualization

Travel patterns can be visualized using `GPE_plot_travel()`. 

The function takes a visits data frame that includes `time_minutes` and `distance_km` columns (i.e. as a result of using `GPE_travel_time_dist()`. By default, it show an histogram with the distribution of travel time (in minutes):

```{r}
GPE_plot_travel(visits_timedist)
```

Setting the parameter `metric = "time"` can be used  to used to show travel distance (in kilometers) instead of time:

```{r}
GPE_plot_travel(visits_timedist, metric = "distance")
```

Instead of an histogram, a [boxplot](https://en.wikipedia.org/wiki/Box_plot) can be obtained by setting the parameter `plot_type = "boxplot"`:

```{r}
GPE_plot_travel(visits_timedist, plot_type = "boxplot")
```

Optionally, an additional data frame can be used as input to merge it with the travel data and plot any of its attributes. This data frame must be contain either participants a locations; that is, besides the attribute to plot it must contain either a "location_id" or a "participant_id" column. When an additional dataframe is used, also specify the name of the column to plot. 

For example, to show travel time by participant income level:

```{r}
GPE_plot_travel(visits_timedist, add_data_from = participants, plot_attribute = "income_bracket")
```

To show travel distance by location type, with boxplots:

```{r}
GPE_plot_travel(visits_timedist, 
                add_data_from = locations, plot_attribute = "type", 
                plot_type = "boxplot", metric = "distance")
```

