---
title: "Using GPE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using GPE}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Prerequisites

You'll need tabular data with addresses, or "lon" and "lat" columns containing WG84 (Mercator) location coordinates. IE, a table with the addresses of schools were program beneficiaries attend, or a street or centroid identifying their neighboorhoods, or their home addresses -if using personal data is allowed.

## Example

Having a "participants" dataframe such as:

```{r}
participants <- read.csv("../data/participants.csv")

head(participants)
```

And another dataframe containing the Program sites, such as:

```{r}
locations <- read.csv("../data/locations.csv")

head(locations)
```

Then the geographic position of participants and program locations can be plotted, on top of a basemap, with:

```{r}
library(GPE)

GPE_plot_map(participants, locations)
```

