---
title: "Using GPE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using GPE}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, dpi = 300)
```

## Prerequisites

You'll need tabular data with addresses, or "lon" and "lat" columns containing WG84 (Mercator) location coordinates. IE, a table with the addresses of schools were program beneficiaries attend, or a street or centroid identifying their neighboorhoods, or their home addresses -if using personal data is allowed.

## Example


```{r}
library(GPE)
```


You'll need a data frame containg participant data. 

GPE includes "participants", an example data frame with fictional public program beneficiaries:

```{r}
head(participants)
```

Then another  dataframe containing the Program sites -that is, places that participants visit in order to interact with the program, such as training centers, day care provdiders, etc.  

GPE includes "locations", an example data frame with fictional public program sites:


```{r}
head(locations)
```

Then the geographic position of participants and program locations can be plotted, on top of a basemap, with:

```{r}
GPE_plot_map(participants, locations)
```

